---
title: "Project 1 - Data Visualization"
author: "Asher Dvir-Djerassi"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

#Turn off scientific notation 
options(scipen = 999)
#options(scipen = 0) #If you want to turn scientific notation on again
```

And lets preview this data:

```{r}
head(inc)
```

```{r}
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

Distribution and standard deviation of revenue
```{r}
summary(round(inc$Revenue))
sd(inc$Revenue)
```

Distribution and standard deviation of employees
```{r}
summary(na.omit(round(inc$Employees)))
sd(na.omit(inc$Employees))
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (i.e. how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (i.e. taller than wide), which should further guide your layout choices.

```{r}
library(dplyr)
library(ggplot2)

count(inc, State) %>%
  arrange(n) %>%
  mutate(State = factor(State, State)) %>%
  ggplot(aes(x=State, y=n)) + 
  geom_bar(stat="identity") + 
  coord_flip() +
  ylab("Number of Companies") + 
  xlab("States")
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

### For this question, I create a box plot for each industry in New York, the state with 3rd highest number of registered companies. The first box plot visualized is not intellible because of the dominance of outliers. The second box plot zooms in, thereby excluding those outliers from the visual. 

```{r}
#State with 3rd most companies (i.e. New York)
library(dplyr)

#Mean by industry
df <- inc %>% 
  group_by(Industry) %>% 
  filter(State == "NY") %>% 
  summarize(Mean = mean(Employees, na.rm=TRUE)); df

#Median by industry
df <- inc %>% 
  group_by(Industry) %>% 
  filter(State == "NY") %>% 
  summarize(Median = median(Employees, na.rm=TRUE)); df

#Boxplot of number of employees by industry
ggplot(inc, aes(x=Industry, y=Employees)) + 
  geom_boxplot() + 
  coord_flip()

#Zooming in (not restricting outlier): Boxplot of number of employees by industry. NOTE: I choose to not impose on the data by restricting outlier, rather this just zooms into the previous graph.
ggplot(inc, aes(x=Industry, y=Employees)) + 
  geom_boxplot() + 
  coord_flip(ylim = c(0,1000))
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
df <- aggregate(inc[c(4,6)], by = list(inc$Industry), FUN = sum, na.rm=TRUE)

df$per_capita <- df$Revenue/df$Employees

ggplot(df, aes((x=reorder(Group.1, per_capita)), per_capita, group = Group.1, fill = Group.1)) +
  geom_bar(stat="identity")+
  theme(legend.position="none") +
  coord_flip()  +
  ylab("Revenue Per Employee") + 
  xlab("Industry") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

